\section{Introduction}

In case you have not noticed, 3D printers are becoming as common as the X-Acto
knife in home shops these days. They let you, or your kids, build some amazing
parts, some of which find their way into our model airplanes. If you look
around on {\it Thingiverse} \cite{thingy} you will discover that many of the
files needed for 3D printing were generated using a neat open-source {\it
Computer Aided Design} tool called \osc. I have used this tool for a number of
projects, and I decided to see how it could help with my current indoor model
building. In this article I will show how I developed a new design for a \LPP\ model
using \osc. I named the design \MM\ since I used a fair amount of math in its
design.

\osc\ is not your usual CAD tool. It is a programmer's tool, meaning that you
write fairly simple program code that describes your model. Then \osc\ processes
that code to generate a visual representation of your design that you can see on
your computer screen. Once you are happy with the design, \osc\ can export your
design in file formats that could be post-processed as part of the path to a 3D
printer. Unfortunately, we probably are not going to see competition ready 3D
printed indoor models anytime soon, so we will not explore 3D printing in this
article.

Do not be frightened off by having to write program code for \osc. Like
anything new, it can be a bit intimidating for beginners, but I have worked on
making the code needed simple enough that even non-technical folks should be
able to generate useful results.

In {\it Part 1} I will explain the basic concepts used in \osc\ to
set up a 3D design. This is not a complete tutorial, but covers enough to help
you understand the rest of the article, In {\it Part 2} we will work through
the design on my \LPP\ model. You will see some code here, but only enough to
explain how I generated the design. Finally, {\it Part 3} will show some
analysis techniques I used to get a weight and balance assessment of my design.
This is always a good thing to do before going flying! In this section, I will
show a bit of Python code. Python is another programming language, commonly
used in introductory programming courses. Kids in elementary school have
managed to get going in programming using Python.

I have tried to keep the computer jargon to a minimum. Instead, I have come up with
a way to generate \osc\ code that uses terms more familiar to model builders.
Hopefully, this article will encourage you to try these techniques when you
design your own model.

All of the code and more detailed documentation on this project is available on
the project website at {\bf https://github.com/rblack42/math-magik}
\cite{blackr}.


